---
title: "R Notebook"
output: html_notebook
---
this notebook is meant to be used on the combined missouri crash data from 2017-2022 with the purpose of creating block group level variables for pedestrian/crash incidents.

```{r}
library(dplyr)
library(ggplot2)
library(naniar)
library(crosstable)
library(stringr)
library(tidyr)
```
```{r}
crash2017 <- read.csv("Crash Data/MO Crash Data_2017.csv")
crash2018 <- read.csv("Crash Data/MO Crash data_2018.csv")
crash2019 <- read.csv("Crash Data/MO Crash Data_2019.csv")
crash2020 <- read.csv("Crash Data/MO Crash Data_2020.csv")
crash2021 <- read.csv("Crash Data/MO Crash Data_2021.csv")
crash2022 <- read.csv("Crash Data/MO Crash Data_2022.csv")
```
```{r}
all_crash <- bind_rows(crash2017, crash2018, crash2019, crash2020, crash2021, crash2022)
```
```{r}
gg_miss_var(all_crash)
```
there is no missing data in this dataset! what a beautiful clean piece of business
```{r}
cities <- table(all_crash$City)
```
question: what places do we remove from this analysis? is kansas data not part of this work, i don't remember?

```{r}
table(all_crash$Type)
```
```{r}
table(all_crash$Severity)
```

likely going to need to remove some of the superfluous information in the address columns so that lauren can map the intersections more successfully, but i'm not 100% sure yet. she is going to test out the geocoding and see what is important to change and what isn't. 
```{r}
all_crash <- str_replace(all_crash$On.Street, "CST ", "")
all_crash <- str_replace(all_crash$At.Street, "CST ", "")

```
